### 2022-01-18 src-latex/parts/40-ner-approaches.tex
review of section 4

% !TeX root = ../main-paper.tex
\section{NER Approaches}
We select twe deep-learning based NER models available in packaged software libraries: SpaCy NLP pipelines and CamemBERT.
In the case of CamemBERT, we also create an additional model pre-trained on a collection of unannotated directories entries extracted with Pero-OCR.
The three models are then fine-tuned for named entity recognition in historical directories following a procedure detailed in sections~\ref{subsection:experiment-1} and ~\ref{subsection:experiment-2}.

\subsubsection{SpaCy NER pipeline}
Spacy is a software library that offers NLP components assembled in modular pipelines specialised by language.
Although BERT is available in the latest version of SpaCy (v3), the pipeline for French does not provide a NER layer.
Hence we rely on SpaCy's ad hoc pipelines trained on French corpora and capable of named entity recognition.
% deep-sequoia and wikiner-fr, capable of named entity recognition.\textit{fr\_core\_news\_lg}\footnote{https://spacy.io/models/fr} trained two corpora in French: deep-sequoia and wikiner-fr, capable of named entity recognition.
The global architecture of these pipelines have not been yet published but is explained by the developers on their website.
Words are first encoded into local context-aware embeddings using a window-based CNN similar to~\cite{collobert2011}.
The decision layer is an adaptation of the transition-based model presented in~\cite{lample2016}.
As words are processed sequentially, their vectors are concatenated with those of the last known entities to encode the nearby predicted semantics.
The classification layer relies on a finite-state machine whose transition probabilities are learned using a multilayer perceptron.
\bertrand{Drop the next sentences if we need space}
In 2018~\cite{won2018} evaluated SpaCy's NER ability to detect place names in five corpora of ancient letters written in English.
They measured an average F1 score of 0.57.
The SpaCy developers claim an accuracy of 0.85 for the English NER pipeline \textit{en\_core\_web\_lg} on the OntoNotes 5.0 corpus\footnote{https://spacy.io/usage/facts-figures}.
%For our experiments, the French model \textit{fr\_core\_news\_lg} is fine-tuned using our ground truth corpus.

\subsubsection{CamemBERT}
Language model based on Transformers, like BERT, have become a new paradigm for NER\cite{li2020}. 
The learned embeddings can be used as distributed representations for input instead of traditional embeddings like Google Word2Vec, and they can be further fine-tuned for NER by adding an additional output layer. 
They can also be pretrained in an unsupervised way on historical texts for domain adaptation.

As our directories are written in French, we chose the language model CamemBERT \cite{martin-etal-2020-camembert}, a Transformer model trained on a French corpus.
\bertrand{WIP}

For our experiments, two models are derived. One is produced by fine-tuning this generic model on our ground truth dataset.
The second is created by adapting the embeddings of this model using 845 000 raw directory entries, randomly selected and extracted with PERO OCR. The base model is pretrained on these texts for two tasks, namely, Next Sentence Prediction (NSP) and Masked Language Model (MLM). Finally, the resulting model is also fine-tuned using our ground truth corpus.

>>>>

% !TeX root = ../main-paper.tex
\section{NER Approaches}
We select two deep-learning based NER models available in packaged software libraries: SpaCy NLP pipelines and CamemBERT.
We also create an additional CamemBERT model pre-trained on a collection of unannotated directories entries extracted with Pero-OCR.
The NER layer of each three models is then fine-tuned on historical directories as detailed in sections~\ref{subsection:experiment-1} and ~\ref{subsection:experiment-2}.

\paragraph{Naming conventions} The short name of each model, indicated in square brackets, is used to refer to it from now on.

\subsubsection{SpaCy NER pipeline~[SpaCy NER]}
Spacy is a software library that offers NLP components assembled in modular pipelines specialised by language.
Although BERT is available in the latest version of SpaCy (v3), the pipeline for French does not provide a NER layer at the time of our experiments (jan. 2022).
Hence we rely on SpaCy's ad hoc pipelines trained on French corpora and capable of named entity recognition.
% deep-sequoia and wikiner-fr, capable of named entity recognition.\textit{fr\_core\_news\_lg}\footnote{https://spacy.io/models/fr} trained two corpora in French: deep-sequoia and wikiner-fr, capable of named entity recognition.
The global architecture of these pipelines have not been yet published but are explained by the developers on their website.
Words are first encoded into local context-aware embeddings using a window-based CNN similar to~\cite{collobert2011}.
The decision layer is an adaptation of the transition-based model presented in~\cite{lample2016}.
As words are processed sequentially, their vectors are concatenated with those of the last known entities to encode the nearby predicted semantics.
The classification layer relies on a finite-state machine whose transition probabilities are learned using a multilayer perceptron.
\bertrand{Drop the next sentences if we need space}
In 2018~\cite{won2018} evaluated SpaCy's NER ability to detect place names in five corpora of ancient letters written in English.
They measured an average F1 score of 0.57.
The SpaCy developers claim an accuracy of 0.85 for the English NER pipeline \textit{en\_core\_web\_lg} on the OntoNotes 5.0 corpus\footnote{https://spacy.io/usage/facts-figures}.
%For our experiments, the French model \textit{fr\_core\_news\_lg} is fine-tuned using our ground truth corpus.

\subsubsection{CamemBERT~[CmBERT]}
Language model based on Transformers, like BERT, have become a new paradigm for NER\cite{li2020}. 
The learned embeddings can be used as distributed representations for input instead of traditional embeddings like Google Word2Vec, and they can be further fine-tuned for NER by adding an additional output layer. 
They can also be pretrained in an unsupervised way on historical texts for domain adaptation.
As the directories are written in French, we chose the language model CamemBERT \cite{martin-etal-2020-camembert}, a Transformer model trained on a French corpus.
\bertrand{WIP}
\subsubsection{CamemBERT pretrained on directories~[CmBERT+ptrn]}
In order to evaluate the benefits of pre-training the language model on OCR texts from domain-related documents, we adapt the CmBERT embeddings using \num{845.000} entries randomly selected wihtin the collection of directories and extracted with Pero-OCR.
We achieve this by training the CmBERT model for 3 epochs on two unsupervised tasks: next sentence prediction and masked language model.

### 2022-01-18 src-latex/parts/50-xp-setup.tex
Add some comments
review of section 4

\Cref{fig.protocol} depicts the evaluation procotol used to assess the OCR and NER systems. Two quality evaluation
are performed in the pipeline named respectively \emph{OCR Q.A.} and \emph{NER Q.A.}.

First the layout extraction and the entry segmentation of the page are performed with a semi-automated system and
checked by a human. Afterward, an OCR system runs on the thumbnails of each segmented entry to extract its text. An
entry might span over multiple text lines but is always single-block. Thus, the most adapted mode is chosen when
the OCR system allows for detection mode (e.g. the \emph{block} mode for \emph{Tesseract}).

\textbf{OCR Q.A.} is performed after some text normalization of the OCR system outputs. Text normalization consists in
projecting unicode characters onto the latin-1 charset (whenever it makes sense) and removing extra symbols (hands,
crosses). Then, the predicted text is aligned with the reference text using standard tools and the Character Error Rates
(CER) are computed at entry level and at global level.

@@

In this benchmark, we consider 3 OCR systems well-known for the historical document analysis: Pero OCR, Tesseract and Kraken.

@@

Next, a NER system extracts the named entities from the text of each entry output by the OCR and the \textbf{NER Q.A.}
is performed. The NER system outputs a text with tags that enclose the entities. To assess the quality of the entity
extraction, we rely on a technique similar as for the OCR evaluation. The predicted text is aligned with the reference
text and the tags are projected in the alignment. Then, the precision, recall and f-measure (or f-score) are computed considering
only the exact matches between entities. Precision and recall are defined by equation (2); the f-measure in their harmonic mean.

@@

The whole process is illustrated on~\cref{fig.eval-ocr-ner}. The OCR text and the GOLD text are first aligned to
evaluate the OCR accuracy. As there are 11 mismatches over 56 aligned characters, the CER is thus 24\%. This alignment
is then used to back-project the GOLD tags to build the tagged NER reference string. Finally, the NER system runs on the
OCR text; its output is compared to the NER reference string. There is only 2 over 3 tags matching in the prediction (precision),
while only 2 over 4 entities are matching in the reference (recall). It follows an overall f-score of 0.4.

@@

\subsection{NER sensibility to the number of training samples}
\label{subsection:experiment-1}
The first experiment evaluates the performances of the three models on training sets of different sizes.
To do so, we split the gold reference into a training set, a development set, and a test set. The training set is then gradually reduced in size while maintaining the relative frequency of directories within.
The training and testing procedure is the same for the three models.

@@

To generate smaller training sets we start from the initial training set and iteratively split it in half using the same stratified sampling strategy as before.
We stop if there is only one entry left in a directory or if the current train set contains less than 30 entries.
Applying this procedure to the initial training set produced 8 training sets containing 49, 99, 199, 398, 796, 1593, 3186 and 6373 entries.

@@

\subsection{NER in noisy OCRed texts}
\label{subsection:experiment-2}
The second experiment assesses the impact of the OCR used on the quality of the NER results. To this end, we use directory entries whose text has been extracted, corrected and annotated manually (see Section 3). This reference text is aligned with the text produced on the same entries by Tesseract and PERO OCR. The named entities annotated in the reference text are then projected into the OCRed texts. To make sure that all training, development and test datasets contain the same annotations, only the entries for which the same number of annotations are found in the reference text and in the OCRised texts are kept. Finally, the reference text is divided into three datasets, using the same method as in Experiment 1. Two additional test files are generated with exactly the same entries as the test file produced from the reference text: one from the text produced by PERO OCR and one from that produced by Tesseract.

@@

\subsubsection{OCR systems parameters}
\nathalie{Y a-t-il des précisions à ajouter sur la façon dont ont été utilisés les OCR? Ou bien ce sont les versions "sur étagère" qui ont servi?}
\begin{itemize}
    \item Tesseract 4
    \item Pero
\end{itemize}

\subsubsection{Spacy}
We used the French model \textit{fr\_core\_news\_lg} provided by Spacy\mcite{spacy} library and leveraged the library utilities to fine-tune it with various training and validation datasets derived from our ground truth dataset (see section 5.2 for details). The base model is fine-tuned in turn with each of the 8 generated training data sets, using the same training parameters. Early stopping is activated with patience parameters set at 1600. For each size of training dataset, 5 runs are performed, always with the same parameters, and the mean value of each performance score is computed.

@@

\subsubsection{Huggingface CamemBERT}
We started from the CamemBERT-ner model already fine-tuned on WikiNER-fr corpus and published on the Hugging Face repository \footnote{\url{https://huggingface.co/Jean-Baptiste/camembert-ner}}. Its output layer is a linear model with a Softmax function. Two NER models are derived from this base model.

\textcolor{green}{Ex 2: On garde CamemBERT et CamemBERT pretrained fine-tuned sur le train clean (patience = 5), validé avec dev clean. Teste les deux modèles sur les trois datasets de test.}

The base model is pretrained on 845 000 entries extracted with PERO OCR for the Next  Sentence  Prediction  (NSP) and  Masked  Language  Model(MLM) tasks. Then it is fine-tuned with the same parameters and on the same training datasets as the first model.

>>>>

The Figure \Cref{fig.protocol} depicts the evaluation protocol used to assess the OCR and NER systems. Two quality evaluations
are performed in the pipeline, respectively named \emph{OCR Q.A.} and \emph{NER Q.A.}.

First, the layout extraction and entry segmentation of the page are performed with a semi-automated system and
checked by a human. Afterward, an OCR system runs on the thumbnails of each segmented entry to extract its text. An
entry might span over multiple text lines but is always a single block. Thus, the most adapted mode is chosen when
the OCR system allows for the detection mode (e.g. the \emph{block} mode for \emph{Tesseract}).

\textbf{OCR Q.A.} is performed after some text normalization of the OCR system outputs. Text normalization consists in
projecting Unicode characters onto the latin-1 charset (whenever it makes sense) and removing extra symbols (hands,
crosses). Then, the predicted text is aligned with the reference text using standard tools and the Character Error Rates
(CER) are computed at the entry level and at the global level.

@@

In this benchmark, we consider 3 OCR systems well known for historical document analysis: Pero OCR, Tesseract, and Kraken.

@@

Next, a NER system extracts the named entities from the text of each entry output by the OCR and the \textbf{NER Q.A.}
is performed. The NER system outputs a text with tags that enclose the entities. To assess the quality of the entity
extraction, we rely on a technique similar as for the OCR evaluation. The predicted text is aligned with the reference
text and the tags are projected in the alignment. Then, the precision, recall, and f-measure (or f-score) are computed considering
only the exact matches between entities. Precision and recall are defined by equation (2); the f-measure is their harmonic mean.

@@

The whole process is illustrated on~\cref{fig.eval-ocr-ner}. The OCR text and the GOLD text are first aligned to
evaluate the OCR accuracy. As there are 11 mismatches over 56 aligned characters, the CER is thus 24\%. This alignment
is then used to back-project the GOLD tags to build the tagged NER reference string. Finally, the NER system runs on the
OCR text; its output is compared to the NER reference string. There is only 2 over 3 tags matching in the prediction (precision),
while only 2 over 4 entities are matched in the reference (recall). It follows an overall f-score of 0.4.

@@

\subsection{Experiment 1: NER sensibility to the number of training examples}
\label{subsection:experiment-1}
The first experiment evaluates the performances of the three models on training sets of different sizes.
To do so, we split the gold reference into a training set, a development set, and a test set. The training set is then gradually reduced in size while maintaining the relative frequency of directories within.
The training and testing procedure is the same for the three models.

@@

To generate smaller training sets, we start from the initial training set and iteratively split it in half using the same stratified sampling strategy as before.
We stop if there is only one entry left in a directory or if the current train set contains less than 30 entries.
Applying this procedure to the initial training set produced 8 training sets containing 49, 99, 199, 398, 796, 1593, 3186, and 6373 entries.

@@

\subsection{Experiment 1: NER on texts with OCR noise}
\label{subsection:experiment-2}
The second experiment assesses the impact of the OCR on the quality of the NER results. To this end, we use directory entries whose text has been extracted, corrected, and annotated manually (see Section 3). This reference text is aligned with the text produced on the same entries by Tesseract and PERO OCR. The named entities annotated in the reference text are then projected into the OCRed texts. To make sure that all training, development, and test datasets contain the same annotations, only the entries for which the same number of annotations are found in the reference text and in the OCRised texts are kept. Finally, the reference text and the PERO OCR text are divided into three datasets, using the same method as in Experiment 1. An additional test file is generated from the text produced by Tesseract with the same entries as the test files produced from the reference and PERO OCR texts.

@@

\subsubsection{OCR system parameters}
\nathalie{Y a-t-il des précisions à ajouter sur la façon dont ont été utilisés les OCR? Ou bien ce sont les versions "sur étagère" qui ont servi?}
\begin{itemize}
    \item Tesseract 4
    \item Pero
\end{itemize}

\subsubsection{Spacy}
We use the French model \textit{fr\_core\_news\_lg} provided by Spacy\mcite{spacy} library and leverage the library utilities to fine-tune it on various training and validation datasets derived from our ground truth dataset (see Section 5.2 for details). The base model is fine-tuned in turn on each of the 8 generated training datasets, using the same training parameters. Early stopping is activated with patience parameters set at 1600. For each size of the training dataset, 5 runs are performed, always with the same parameters, and the mean value of each performance score is computed.

@@

\subsubsection{Huggingface CamemBERT}
We start from the CamemBERT-ner model already fine-tuned on WikiNER-fr corpus and published on the Hugging Face repository \footnote{\url{https://huggingface.co/Jean-Baptiste/camembert-ner}}. Its output layer is a linear model with a Softmax function. Four NER models are derived from this base model:

$BERT_{reference}$ The base model is fine-tuned in turn on each of the training datasets generated from the groundtruth corpus. Training parameters are set depending on the considered experiment: early stopping is activated with the patience parameter set at 3 for experiment 1 and at 5 for experiment 2.

$BERT_{pero-ocr}$ This model is used for experiment 2 only. The base model is fine-tuned on the training dataset generated by projecting the named entity annotations on the text extracted by PERO OCR, using the same parameters as for the $BERT_{reference}$ model.

$BERT_{ptrn-reference}$ The base model is pretrained using 845000 entries extracted by PERO OCR with the Next  Sentence  Prediction  (NSP) and  Masked  Language  Model(MLM) tasks. Then it is fine-tuned with the same parameters and on the same training datasets generated from the groundtruth corpus as the $BERT_{reference}$ model.

$BERT_{ptrn-pero-ocr}$ This model is used for experiment 2 only. The base model is pretrained like the $BERT_{ptrn-reference}$ model and it is fine-tuned with the same parameters and on the same training datasets as the $BERT_{pero-ocr}$ model.
